Turbulence in compact to giant H~II regions

### 2023-06-02 strucfunc_paper_plus.tex
Finish and submit the arxiv version of the paper

The Orion Nebula (M42) is the prototype of a galactic \hii{} region located at a distance of 440 pc \citetext{\SI{1}{\arcsecond} = \SI{0.002}{pc}; \citealp{2008AJ....136.1566O}}.
The O7 V star \(\uptheta^{1}\) Ori C is the most luminous and hottest star with an ionizing flux \(Q_0\) of \SI{e49.12}{photons.s^{-1}} \citep{2006A&A...448..351S},
\startNEW
dominating the ionization of the nebula, which has a
\stopNEW
luminosity in \ha{} of \SI{e37.18}{erg.s^{-1}}.
With 3 other B-type stars they form the Trapezium cluster.
The bright region that surrounds these massive ionizing stars is called the Huygens region.
%In the southwest direction of the Huygens region there is a drop in the brightness as one approach a much larger region called Extended Orion Nebula \citetext{EON henceforth;  \citealp{2008Sci...319..309G}}
The position of the Orion Nebula \NEW{on the near side of} the Orion Molecular Cloud (OMC-1)
\startNEW
makes it an ongoing site of star formation with a large population of young stars,
\stopNEW
some of which are sources of stellar jets and Herbig-Haro objects \citep{1993ApJ...410..696O}.
Its physical properties and kinematics, along with its stellar population is documented in \citet{2001ARA&A..39...99O}.

@@

The Lagoon Nebula (M8, NGC~6523) is located at a distance of \SI{1250}{pc}  \citetext{\SI{1}{\arcsecond} = \SI{0.006}{pc} ; \citealp{2005A&A...430..941P}}.
It contains the young stellar cluster NGC~6530 and the region is illuminated by different massive stars of spectral types O and B, the hottest being 9~Sgr, type O4V \citep{Damiani:2017b}.
\startNEW
The optically brightest part of the nebula is denoted the ``Hourglass nebula'',
which is illuminated by the O7 star Herschel~36 \citep{1986AJ.....91..870W}.
\stopNEW
These stars gives an ionizing flux \(Q_0 = \SI{e49.41}{photons.s^{-1}}\) (2 times higher than in the Orion Nebula) and the luminosity in \ha{} is \SI{e37.47}{erg.s^{-1}} \citep{1984ApJ...287..116K}.
The mean non-thermal linewidth in \ha{} is \SI{11.7 \pm 1.6}{km.s^{-1}} \citep{1973ApJ...183..851B}.
\startNEW
Molecular gas and ongoing star formation are present in the region around
Herschel~36 \citep{Arias:2006e, Tiwari:2018a} and the lower-luminosity region
M8~East \citep{1984ApJ...278..170S, Tiwari:2020a}.
\stopNEW
The properties of the nebula are reviewed by \citet{2008hsf2.book..533T}.

@@

The Carina nebula is a large star-forming complex located at a distance of \SI{2 350}{pc} \citetext{\SI{1}{\arcsecond} = \SI{0.01}{pc}; \citealp{2006ApJ...644.1151S}}.
It contains young clusters like Trumpler~14 and~16, and Collinder~228 from the Car~OB1 association, one of the largest OB associations in the Galaxy.
This OB association has more than 65 stars earlier than B0, and several Wolf-Rayet (WR) stars and the most luminous star is the blue variable $\eta$~Car.
The ionizing flux \(Q_0\) is \SI{e50.95}{photons.s^{-1}} \citep{2008hsf2.book..138S} (60 times higher compared to the Orion Nebula) and the luminosity in \ha{} of \SI{e39.01}{erg.s^{-1}}.
As a high-mass star formation region, it presents several phenomena related to young stars,
such as evaporating protoplanetary discs, erosion of large dust pillars, and the triggering of a second generation of embedded stars and Herbig-Haro objects \citetext{see \citealp{2008hsf2.book..138S} and reference therein}.
\startNEW
This region is at an intermediate scale between nearby regions like Orion
that are dominated by a single star, and giant star formation regions such as 30~Doradus.
\stopNEW

@@

The giant \hii{} region 30~Doradus (Tarantula Nebula) is located at a distance of \SI{49.9}{kpc} \citetext{\SI{1}{\arcsecond} = \SI{0.24}{pc}; \citealp{2013Natur.495...76P}}.
It is a luminous star-forming region in the LMC (Large Magellanic Cloud) and is the most luminous complex in the Local Group \citep{1984ApJ...287..116K}.
\startNEW
The star cluster that energizes the region is NGC~2070, with a central core (known as R136)
with ionizing luminosity \(Q_0 = \SI{e51.40}{photons.s^{-1}}\)  \citep{2020MNRAS.499.1918B},
which produces an \ha{} luminosity of \SI{e39.46}{erg.s^{-1}} (200 times the Orion Nebula)
for the central ionized region of size \SI{40}{pc}.
\stopNEW
The mean non-thermal linewidth in \ha{} is \SI{22}{km.s^{-1}} \citep{2013A&A...555A..60T}.
\startNEW
Star formation is still ongoing in NGC~2070 \citep{2013AJ....145...98W},
although the star-formation rate peaked about \SI{4}{Myr} ago
\citep{2015ApJ...811...76C} and it currently hosts some of the most massive stars known
\citep{Kalari:2022p}
and several Wolf-Rayet stars \citep{2011A&A...530A.108E}.
\stopNEW
This nebula is commonly used as a comparison for extragalactic star-forming regions.

@@

NGC 346 is the most active star-formation region in the SMC (Small Magellanic Cloud) located at a distance of 62 kpc \citetext{\SI{1}{\arcsecond} = \SI{0.30}{pc} ; \citealp{2001ApJ...562..303D}}. 
It has more than 30 O stars that ionize N66, the largest \hii{} region \citep{2011ApJ...740...10D}.
\NEW{The cluster}
has an ionizing output \(Q_0\) of \SI{e50.71}{photons.s^{-1}} (40 times higher than the Orion Nebula) and \NEW{the \hii{} region has} an \ha{} luminosity of \SI{e38.77}{erg.s^{-1}} \citep{2010A&A...517A..39H,1984ApJ...287..116K}.
\startNEW
A supernova remnant is present in the region \citetext{J59.4\(-7210\), \citealp{Ye:1991d}}
but it does not overlap with the MUSE field that we observe here \citep{Maggi:2019q}.
High-resolution \ha{} spectroscopy of the \hii{} region
gives a non-thermal velocity width of \(\siglosm = \SI{10.1}{km.s^{-1}}\)
\citep{2003ApJ...586.1179D}.
\stopNEW
\citet{2008ApJ...688.1050G} propose an expanding \hii{} region or bubble blown by the winds of the massive progenitor as a mechanism that shapes the recent star formation in this region, in addition to the photoionization from the OB stars. 
This mechanism is similar to shell-like \hii{} regions, with a central cluster in a cavity, and with ongoing star formation triggered around their periphery.

@@

Hubble V is located near the ``Hodge OB 8'' OB association with a number of \num{\sim 36} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O3 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.23}{photons.s^{-1}} (13 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.30}{erg.s^{-1}}\citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} \(\siglosm = \SI{10.3 \pm 0.03}{km.s^{-1}}\) \citep{1986A&A...160..374H}.

Hubble X is located near the Hodge ``OB 13'' OB association with a number of \num{\sim 29} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O4 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.14}{photons.s^{-1}} (10 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.21}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \(\siglosm = \SI{10.5 \pm 0.02}{km.s^{-1}}\)  \citep{1986A&A...160..374H}.
\citet{1993PASJ...45..693T} studies the kinematics of both regions.

@@

NGC 595 is the second most luminous giant \hii{} region in the M33 galaxy,
\startNEW
excited by a stellar population comprising
250 OB stars, 13 hot supergiants, and 11 Wolf-Rayet stars \citep{1993AJ....105.1400D, 1996AJ....111.1128M}.
\stopNEW
The ionizing output \(Q_0\) is \SI{e50.88}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and
the \hii{} region has a luminosity in \ha{} of \SI{e38.95}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{16.8 \pm 0.1}{km.s^{-1}}\) \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find an \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the
\startNEW
most luminous \hii{} region in M33 and
\stopNEW
the second most luminous in the Local Group after 30~Dor.
\startNEW
The dominant stellar cluster is located in the center of the region and contains 200 OB stars and 4 Wolf-Rayet stars \citep{1996ApJ...456..174H, 2011MNRAS.411..235E}.
The optically visible stars come from a population
with an age of \SI{4}{Myr},
although infrared observations detect the presence of
ongoing massive star formation in the region
\citep{2012AJ....143...43F, 2012ApJ...761....3M}.
\stopNEW
The ionizing output is \(Q_0 = \SI{e51.35}{photons.s^{-1}}\)
(175 times higher than the Orion Nebula) and
the region has an \ha{} luminosity of \SI{e39.42}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{17.8 \pm 0.3}{km.s^{-1}}\) \citep{1986A&A...160..374H}.
\startNEW
The morphology of the region is strikingly similar to that of 30~Dor
(see Figure~\ref{fig:hii-regions})
\stopNEW
and is visually dominated by a predominant big loop, many expanding shells and different size filaments.

@@

\section{Additional covariance corner plots for model fits}
\label{sec:addit-covar-corn}
Supplementary material
\href{https://github.com/will-henney/javier-strucfunc-paper/blob/main/supplementary_material_strucfunc_paper.pdf}
{available online}.

>>>>

The Orion Nebula (M42) is the prototype of a galactic \hii{} region located at a distance of 440 pc \citetext{\SI{1}{\arcsecond} = \SI{0.002}{pc}; \citealp{2008AJ....136.1566O}}.
The O7 V star \(\uptheta^{1}\) Ori C is the most luminous and hottest star with an ionizing flux \(Q_0\) of \SI{1.3e49}{photons.s^{-1}} \citep{2006A&A...448..351S},
\startNEW
dominating the ionization of the nebula, which has a
\stopNEW
luminosity in \ha{} of \SI{1.5e37}{erg.s^{-1}}.
With 3 other B-type stars they form the Trapezium cluster.
The bright region that surrounds these massive ionizing stars is called the Huygens region.
%In the southwest direction of the Huygens region there is a drop in the brightness as one approach a much larger region called Extended Orion Nebula \citetext{EON henceforth;  \citealp{2008Sci...319..309G}}
The position of the Orion Nebula \NEW{on the near side of} the Orion Molecular Cloud (OMC-1)
\startNEW
makes it an ongoing site of star formation with a large population of young stars,
\stopNEW
some of which are sources of stellar jets and Herbig-Haro objects \citep{1993ApJ...410..696O}.
Its physical properties and kinematics, along with its stellar population is documented in \citet{2001ARA&A..39...99O}.

@@

The Lagoon Nebula (M8, NGC~6523) is located at a distance of \SI{1250}{pc}  \citetext{\SI{1}{\arcsecond} = \SI{0.006}{pc} ; \citealp{2005A&A...430..941P}}.
It contains the young stellar cluster NGC~6530 and the region is illuminated by different massive stars of spectral types O and B, the hottest being 9~Sgr, type O4V \citep{Damiani:2017b}.
\startNEW
The optically brightest part of the nebula is denoted the ``Hourglass nebula'',
which is illuminated by the O7 star Herschel~36 \citep{1986AJ.....91..870W}.
\stopNEW
These stars gives an ionizing flux \(Q_0 = \SI{2.6e49}{photons.s^{-1}}\) (2 times higher than in the Orion Nebula) and the luminosity in \ha{} is \SI{2.95e37}{erg.s^{-1}} \citep{1984ApJ...287..116K}.
The mean non-thermal linewidth in \ha{} is \SI{11.7 \pm 1.6}{km.s^{-1}} \citep{1973ApJ...183..851B}.
\startNEW
Molecular gas and ongoing star formation are present in the region around
Herschel~36 \citep{Arias:2006e, Tiwari:2018a} and the lower-luminosity region
M8~East \citep{1984ApJ...278..170S, Tiwari:2020a}.
\stopNEW
The properties of the nebula are reviewed by \citet{2008hsf2.book..533T}.

@@

The Carina nebula is a large star-forming complex located at a distance of \SI{2 350}{pc} \citetext{\SI{1}{\arcsecond} = \SI{0.01}{pc}; \citealp{2006ApJ...644.1151S}}.
It contains young clusters like Trumpler~14 and~16, and Collinder~228 from the Car~OB1 association, one of the largest OB associations in the Galaxy.
This OB association has more than 65 stars earlier than B0, and several Wolf-Rayet (WR) stars and the most luminous star is the blue variable $\eta$~Car.
The ionizing flux \(Q_0\) is \SI{8.9e50}{photons.s^{-1}} \citep{2008hsf2.book..138S} (60 times higher compared to the Orion Nebula) and the luminosity in \ha{} of \SI{e39}{erg.s^{-1}}.
As a high-mass star formation region, it presents several phenomena related to young stars,
such as evaporating protoplanetary discs, erosion of large dust pillars, and the triggering of a second generation of embedded stars and Herbig-Haro objects \citetext{see \citealp{2008hsf2.book..138S} and reference therein}.
\startNEW
This region is at an intermediate scale between nearby regions like Orion
that are dominated by a single star, and giant star formation regions such as 30~Doradus.
\stopNEW

@@

The giant \hii{} region 30~Doradus (Tarantula Nebula) is located at a distance of \SI{49.9}{kpc} \citetext{\SI{1}{\arcsecond} = \SI{0.24}{pc}; \citealp{2013Natur.495...76P}}.
It is a luminous star-forming region in the LMC (Large Magellanic Cloud) and is the most luminous complex in the Local Group \citep{1984ApJ...287..116K}.
\startNEW
The star cluster that energizes the region is NGC~2070, with a central core (known as R136)
with ionizing luminosity \(Q_0 = \SI{2.5e51}{photons.s^{-1}}\)  \citep{2020MNRAS.499.1918B},
which produces an \ha{} luminosity of \SI{2.9e39}{erg.s^{-1}} (200 times the Orion Nebula)
for the central ionized region of size \SI{40}{pc}.
\stopNEW
The mean non-thermal linewidth in \ha{} is \SI{22}{km.s^{-1}} \citep{2013A&A...555A..60T}.
\startNEW
Star formation is still ongoing in NGC~2070 \citep{2013AJ....145...98W},
although the star-formation rate peaked about \SI{4}{Myr} ago
\citep{2015ApJ...811...76C} and it currently hosts some of the most massive stars known
\citep{Kalari:2022p}
and several Wolf-Rayet stars \citep{2011A&A...530A.108E}.
\stopNEW
This nebula is commonly used as a comparison for extragalactic star-forming regions.

@@

NGC 346 is the most active star-formation region in the SMC (Small Magellanic Cloud) located at a distance of 62 kpc \citetext{\SI{1}{\arcsecond} = \SI{0.30}{pc} ; \citealp{2001ApJ...562..303D}}. 
It has more than 30 O stars that ionize N66, the largest \hii{} region \citep{2011ApJ...740...10D}.
\NEW{The cluster}
has an ionizing output \(Q_0\) of \SI{5.1e50}{photons.s^{-1}} (40 times higher than the Orion Nebula) and \NEW{the \hii{} region has} an \ha{} luminosity of \SI{5.9e38}{erg.s^{-1}} \citep{2010A&A...517A..39H,1984ApJ...287..116K}.
\startNEW
A supernova remnant is present in the region \citetext{J59.4\(-7210\), \citealp{Ye:1991d}}
but it does not overlap with the MUSE field that we observe here \citep{Maggi:2019q}.
High-resolution \ha{} spectroscopy of the \hii{} region
gives a non-thermal velocity width of \(\siglosm = \SI{10.1}{km.s^{-1}}\)
\citep{2003ApJ...586.1179D}.
\stopNEW
\citet{2008ApJ...688.1050G} propose an expanding \hii{} region or bubble blown by the winds of the massive progenitor as a mechanism that shapes the recent star formation in this region, in addition to the photoionization from the OB stars. 
This mechanism is similar to shell-like \hii{} regions, with a central cluster in a cavity, and with ongoing star formation triggered around their periphery.

@@

Hubble V is located near the ``Hodge OB 8'' OB association with a number of \num{\sim 36} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O3 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{1.7e50}{photons.s^{-1}} (13 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{2e38}{erg.s^{-1}}\citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} \(\siglosm = \SI{10.3 \pm 0.03}{km.s^{-1}}\) \citep{1986A&A...160..374H}.

Hubble X is located near the Hodge ``OB 13'' OB association with a number of \num{\sim 29} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O4 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{1.4e50}{photons.s^{-1}} (10 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{1.6e38}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \(\siglosm = \SI{10.5 \pm 0.02}{km.s^{-1}}\)  \citep{1986A&A...160..374H}.
\citet{1993PASJ...45..693T} studies the kinematics of both regions.

@@

NGC 595 is the second most luminous giant \hii{} region in the M33 galaxy,
\startNEW
excited by a stellar population comprising
250 OB stars, 13 hot supergiants, and 11 Wolf-Rayet stars \citep{1993AJ....105.1400D, 1996AJ....111.1128M}.
\stopNEW
The ionizing output \(Q_0\) is \SI{7.6e50}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and
the \hii{} region has a luminosity in \ha{} of \SI{8.9e38}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{16.8 \pm 0.1}{km.s^{-1}}\) \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find an \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the
\startNEW
most luminous \hii{} region in M33 and
\stopNEW
the second most luminous in the Local Group after 30~Dor.
\startNEW
The dominant stellar cluster is located in the center of the region and contains 200 OB stars and 4 Wolf-Rayet stars \citep{1996ApJ...456..174H, 2011MNRAS.411..235E}.
The optically visible stars come from a population
with an age of \SI{4}{Myr},
although infrared observations detect the presence of
ongoing massive star formation in the region
\citep{2012AJ....143...43F, 2012ApJ...761....3M}.
\stopNEW
The ionizing output is \(Q_0 = \SI{2.2e51}{photons.s^{-1}}\)
(175 times higher than the Orion Nebula) and
the region has an \ha{} luminosity of \SI{2.6e39}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{17.8 \pm 0.3}{km.s^{-1}}\) \citep{1986A&A...160..374H}.
\startNEW
The morphology of the region is strikingly similar to that of 30~Dor
(see Figure~\ref{fig:hii-regions})
\stopNEW
and is visually dominated by a predominant big loop, many expanding shells and different size filaments.

@@

% ArXiv version has supplementary material included in same file
\input{sf_appendix_D}

% \section{Additional covariance corner plots for model fits}
% \label{sec:addit-covar-corn}
% Supplementary material
% \href{https://github.com/will-henney/javier-strucfunc-paper/blob/main/supplementary_material_strucfunc_paper.pdf}
% {available online}.
